---
title: "Activity Monitoring Data"
output: 
  html_document: 
    keep_md: yes
---


```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(extrafont)
library(reshape2)
library(plyr)
library(RColorBrewer)
library(lubridate)
library(lattice)
#loadfonts(device = "win")
#########################
# SWITCH BETWEEN LAPTOP AND PC
# PREFIX LAPTOP
#
# prefix <- "C:/Users/Sebastian/Documents/Projects/MBA/50 - BMP/85 - CCB/02 - Graphics/"
#
# PREFIX PC
#
prefix  <- "E:/Dokumente/Projects/DataScience/Coursera Specialization/05 - Reproducible Research/code"
#
#########################
wd <- prefix#paste(prefix, "test", sep = "")

```
## Switching to working directory and loading the csv
We set the working directory.
We read in the activity.csv.
The date column has to be transformed to the Date type.
```{r, echo=TRUE, warning=FALSE, message=FALSE}
setwd(wd)
df <- read.csv("activity.csv", sep = ",", header = TRUE, na.strings = "NA")
df$date <- as.Date(df$date, "%Y-%m-%d")
```
Calculate the sum of the steps per day. All log entries from the same day
are summed up. 
```{r, echo=TRUE, warning=FALSE, message=FALSE}
steps_per_day <- tapply(df$steps, format(df$date, '%Y-%m-%d'), sum) 
```
## Histogram of the steps per day
In this section we show the histogram of the steps logged per day. 
```{r, echo=TRUE, warning=FALSE, message=FALSE}
qplot(steps_per_day, geom="histogram") 
```

```{r, echo=TRUE, warning=FALSE, message=FALSE}
# Calculate mean
steps_mean <- mean(steps_per_day, na.rm = TRUE)
print(steps_mean)
# Calculate median
steps_median <- median(steps_per_day, na.rm = TRUE)
print(steps_median)
```

```{r, echo=TRUE, warning=FALSE, message=FALSE}
# Calculate steps per interval
df_cmp <- df[complete.cases(df),]
steps_per_interval <- tapply(df_cmp$steps, df_cmp$interval, mean)
steps_per_interval <- aggregate(df_cmp$steps, by=list(df_cmp$interval), mean)


```
## Plot time series of steps per interval
In the following figure I show the average of steps per 5 minutes
interval across all dates.
```{r, echo=TRUE, warning=FALSE, message=FALSE}
names(steps_per_interval)[names(steps_per_interval) == 'Group.1'] <- 'Interval'
g <- ggplot(steps_per_interval, aes(x = Interval, y = x))
g2 <- g + geom_line(color = "steelblue", size = 1)
print(g2)
```

## Maximum number of steps in interval

In the following I determine the interval with the maximum number of steps:

```{r, echo=TRUE, warning=FALSE, message=FALSE}
stp_int_ord <- steps_per_interval[with(steps_per_interval, order(-x)), ]
print(stp_int_ord[1,1])
```

## Count the numbers of rows containing NA
```{r, echo=TRUE, warning=FALSE, message=FALSE}
sum(is.na(df))

```
## Fill in NAs with generated values
To fill the NAs with other values I use the following strategy:

If an NA value is found the interval it occurs in is replaced by the mean over the steps per this respective interval over all days.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
# Fill NAs with means from steps_per_interval: 
# 
df_filled <- df
df_filled$steps <- ifelse(is.na(df$steps), steps_per_interval$x[match(df$interval, steps_per_interval$Interval)], df$steps)
```

## Recalculate the steps per day
Here I recalculate the sum of the steps per day with the dataframe in which the NAs have been replaced. Then I show the corresponding histogram of that data and calculate again the mean and the median.
```{r, echo=TRUE, warning=FALSE, message=FALSE}
steps_per_day_filled <- tapply(df_filled$steps, format(df_filled$date, '%Y-%m-%d'), sum) 
qplot(steps_per_day_filled, geom="histogram") 

steps_mean_filled <- mean(steps_per_day_filled, na.rm = TRUE)
print(steps_mean_filled)
steps_median_filled <- median(steps_per_day_filled)
print(steps_median_filled)
```
The values for mean and median do not deviate from the above values very much. The impact of imputing missing values with this strategy is rather low in this case.

# Compare weekdays with weekend
Here I want to compare the profile of steps per interval at weekdays with that during the week. 
```{r, echo=TRUE, warning=FALSE, message=FALSE}

df_filled$we <- ifelse(wday(df_filled$date) > 5, 1, 0)


steps_per_interval_f <- tapply(df_filled$steps, df_filled$interval, mean)
steps_per_interval_f <- aggregate(df_filled$steps, by=c(list(df_filled$interval), list(df_filled$we)), mean)
names(steps_per_interval_f)[names(steps_per_interval_f) == 'Group.1'] <- 'Interval'
names(steps_per_interval_f)[names(steps_per_interval_f) == 'Group.2'] <- 'Weekend'
names(steps_per_interval_f)[names(steps_per_interval_f) == 'x'] <- 'Steps'

steps_per_interval_f$Weekend[steps_per_interval_f$Weekend == 0] <- "No weekend"
steps_per_interval_f$Weekend[steps_per_interval_f$Weekend == 1] <- "Weekend"

pg <- ggplot(steps_per_interval_f, aes(Interval, Steps)) + geom_line(color = "steelblue", size = 1) + geom_point() + 
  facet_wrap(~Weekend, ncol = 1)
```

We can see that activity is more distributed in a homogenous way on the weekend case. Here, the person is more active over the whole day. During the week, a peak occurs in the morning followed by a long period of low activity and a small one at the end of the day.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
print(pg)
```

